### 2018-09-28 paper/preliminaries.tex
couple of small changes

In this section, we present CSP and the concept of minimal hitting sets, which
is central to our novel notion of test for failures refinement.

% =========================================================================
\subsection{CSP, Refinement, and Normalised Transition Graphs}

@@

Throughout this paper, the alphabet of the processes, that is the set of
events that are in scope, is denoted by $\Sigma$ and supposed to be finite.
The FDR tool~\cite{fdr} supports model checking and semantic analyses of
finite-state CSP processes.

@@

Two forms of choice support branching behaviour. An external choice $P
\extchoice Q$ between processes $P$ and $Q$ offers to the environment the
initial events of $P$ and $Q$. Once synchronisation takes place, the process
that has offered this event is chosen and the other is discarded. In an
internal choice $P \intchoice Q$, the environment does not have an
opportunity to interfere:~the choice is made by the process.
%
\begin{example}\label{example:CSP}
  We consider $P$, $Q$, and $R$ below. $P$ is initially
  ready to engage in the event $a$, and then makes an internal  choice to
  behave like either $Q$ or $R$.
  \begin{eqnarray*}
  P & = & a \then (Q\intchoice R)
  \\
  Q & = & a \then P \extchoice c \then P
  \\
  R & = & b \then P \extchoice c \then R
  \end{eqnarray*}
  $Q$, for instance, offers to the environment the choice to engage in $a$
  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
  chosen, the recursion is to $R$.
  \xbox
\end{example}
%
Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
external and internal choice operators allow us to define a choice over a
collection of processes $P(i)$. By convention,  if the index set $I$ is
empty, external choices evaluate to the process $\Stop$, which
deadlocks:~does not engage into any event and does not terminate. An iterated
internal choice is not defined for an empty index set. For an external choice 
indexed over a set of events in the form
$\Extchoice e: A @ e \then P(e), A\subseteq \Sigma$, we use 
abbreviation $e:A \then P(e)$.

@@

Interactions that are not supposed to be visible to the environment can be
hidden. The operator $P \hide H$ defines a process that behaves as $P$, with
the interactions modelled by events in the set $H$ hidden. Frequently, hiding
is used in conjunction with parallelism:~it is often desirable to make
internal actions of each process in a network of parallel processes, perhaps
used for coordination of the network, invisible, while events happening at
their interfaces remain observable.

@@

Failures refinement, $P \lessdet_F Q$, is defined as expected by set
containment $\fails(Q)\subseteq\fails(P)$. Since refusals are subset-closed,
$P \lessdet_F Q$ implies
$(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. So, for
divergences-free processes, failures refinement implies trace refinement.
Therefore, using the conformance relation
%
\begin{equation}\label{eq:conf}
  Q\ conf\ P \defs \forall s\in \trc(P) \cap \trc(Q): \refs(Q/s)
  \subseteq \refs(P/s),
\end{equation}
%
failures refinement can be expressed by $\lessdet_T$ and $conf$.
%
\begin{equation}\label{thm:fref-tref-conf}
(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)
\end{equation}
%

For finite CSP processes, since refusals are
subset-closed, $\refs(P/s)$ can be constructed from the set of \emph{maximal
refusals} defined by
%
\begin{equation}
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
\end{equation}
%
Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
reconstruct the refusals in the set $\refs(P/s)$ as described by
%
\begin{equation}\label{eq:refMaxRef}
\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
\end{equation}
%
Deterministic process states $P/s$ have exactly the one maximal refusal
defined by $\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of
$P/s$, that is, the events that $P/s$ may engage into. The maximal refusals
in combination with the initials of a process express its degree of
nondeterminism.
%
\begin{example}
\label{ex:nondetdegree} $P = (\Stop \intchoice Q)$ has maximal refusals
$\maxrefs(P) = \{ \Sigma \}$, because $\Stop$ refuses to engage in any event,
and this is carried over to $P$ by the internal choice. However, $P$ is
distinguished from $\Stop$ by its initials, which are defined by $[P]^0 = [
\Stop \intchoice Q]^0 = [Q]^0$. So $P$ may engage nondeterministically in any
initial event of $Q$, but also refuse everything, due to internal selection
of $\Stop$.

@@

As shown in~\cite{Roscoe:1994:chapter}, the failures semantics of any
finite-state CSP process $P$ can be represented by a \emph{normalised
transition graph} $G(P)$ defined by a tuple
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
$t$, then there is a transition, that is, an outgoing edge, from $n$ with
label $e$, leading to node $t(n,e)$. Normalisation of $G(P)$ is reflected by
the fact that $t$ is a function. The graph construction
in~\cite{Roscoe:1994:chapter} implies that all nodes $n$ in $N$ are
reachable by sequences of edges labelled by $e_1\dots e_k$ and connecting
states $\ii n,n_1,\dots,n_{k-1},n$, such that
\[
n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
n= t(n_{k-1},e_k).
\]
%
By construction, $s\in\Sigma^*$ is a trace of $P$, if, and only if, there is
a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
$s$. In analogy to $\trc(P)$, we use the notation $\trc(G(P))$ for the set of
finite, initialised paths through $G(P)$, each path represented by its finite
sequence of edge labels. We note that $\trc(P) = \trc(G(P))$. Since $G(P)$ is
normalised, there is a unique node reached by applying the events from $s$
one by one, starting in $\ii n$. Therefore, $G(P)/s$  is also well defined.
By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events occurring
as labels in any outgoing transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
The graph construction from~\cite{Roscoe:1994:chapter} used to define
$G(P)$ for any process $P$ guarantees that $[G(P)/s]^0 = [P/s]^0$ for all
traces $s$ of $P$.

@@

When investigating  tests for failures refinement, the notion of
\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
is also useful. An acceptance set of $P/s$ is a subset of its initials
$[P/s]^0$; equivalently, it is a subset of events labelling  outgoing transitions of
$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
this set. If $P/s$ is nondeterministic, it internally chooses one of its
\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
{\it possibly} refusing the events from $[P/s]^0 - A$ and
{\it always} refusing those from $\Sigma - [P/s]^0$.
The acceptances of $P/s$ are denoted by
$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
the following properties.
%
\begin{eqnarray}
A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
\label{eq:accref}
\\
\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
\label{eq:accinitials}
\\
 X\in\accs(P/s) &\Leftrightarrow & \exists A\in \minaccs(P/s): A\subseteq X \subseteq [P/s]^0
 \label{eq:accsubset}
\end{eqnarray}
%
%Note that (\ref{eq:accref}) can be regarded as a definition of minimal
%acceptances by means of maximal refusals. In this case,
%(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
%definition and the fact that refusals are subset-closed. \fixme{alcc: are you
%sure? (4) does not talk about Acc. Is this important?}
%\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
%removed it.}
%
Exploiting formula (\ref{eq:accref}), 
%(\ref{eq:accinitials}), and (\ref{eq:accsubset}), 
every node of a normalised transition graph can
alternatively be labelled with their minimal acceptances, and this
information is equivalent to that contained in the maximal refusals. Since
process states $P/s$ are equivalently expressed by states $G(P)/s$ of $P$'s
normalised transition graph, we also write $\minaccs(G(P)/s)$ and
note that (\ref{eq:refsAndR}) and (\ref{eq:accref}) imply
\begin{equation}\label{eq:maxrefsminaccs}
\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \}.
\end{equation}
%\fixme{alcc: proof in the report perhaps?}

% .....................................................................................
 \begin{figure}[t]
   %%\hspace*{-40mm}
   \begin{center}
     \includegraphics[width=\textwidth]{p.pdf}
   \end{center}
   %%\vspace*{-10mm}
   \caption{Normalised transition graph of CSP process $P$ from Example~\ref{ex:a}.}
   \label{fig:tga}
 \end{figure}
% .......................................................................................

@@

Having engaged in $c$, the next process state is represented by node 2. Due
to normalisation, there is only a single transition satisfying
$t(1,c) = 2$. This transition, however, can have
been caused by either $Q$ or $R$ engaging into $c$, so node 2 corresponds to
process state $Q/c \intchoice R/c = P \intchoice R$. This is reflected by the
two minimal acceptance sets labelling node 2. Similar considerations explain
the other nodes and transitions in Fig.~\ref{fig:tga}. 
\xbox
\end{example}
%
Summarising, refinement relations between finite-state CSP processes $P, Q$ can be be
expressed by means of their normalised transition graphs
\begin{eqnarray*}
G(P) & = & ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))
\\
G(Q) & = & ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))
\end{eqnarray*}
%
as established by the results  in the following lemma.
There, result (\ref{eq:trcrefa}) reflects trace refinement in terms of graph
traces; (\ref{eq:failconf}) expresses failures refinement in terms of traces
refinement and $conf$; (\ref{eq:failrefa}) states how $conf$ can be expressed
by means of the maximal refusal functions of the graphs involved; and
(\ref{eq:failrefb}) states the same in terms of the minimal acceptances that
can be derived from the maximal refusal functions by means of
(\ref{eq:maxrefsminaccs}). 
%%\fixme{alcc: I think at least the report should point to proofs.}
%
\begin{lemma}
  \label{lemma:tgtrcref}
  \begin{eqnarray}
  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
  \label{eq:trcrefa}
  \\
  \label{eq:failconf}
  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge Q\ conf\ P
  \\
  \label{eq:failrefa}
  Q\ conf\ P & \Leftrightarrow & \nonumber
  \forall s\in\trc(G(Q))\cap \trc(G(P)), R_Q\in r_Q(G(Q)/s):  \nonumber
  \\ & & \tabd
  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
  \\
  \label{eq:failrefb}
   & \Leftrightarrow &
    \forall s\in\trc(G(Q))\cap \trc(G(P)), A_Q\in\minaccs(G(Q)/s):  \nonumber
   \\ & & \tabd
  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
 \end{eqnarray}
  \xbox
\end{lemma}
%
\begin{proof}
Recall that $P\lessdet_T Q$ is defined
as $\trc(Q)\subseteq \trc(P)$ and that $\trc(P)=\trc(G(P))$ and $\trc(Q)=\trc(G(Q))$.
This shows (\ref{eq:trcrefa}).

@@

For proving our main theorems, it is necessary to consider the \emph{product}
of normalised transition graphs. We need this only for the investigation of
corresponding traces in reference processes and processes for SUTs. So, the
labelling of nodes with maximal refusals or minimal acceptances are
disregarded in the product construction. We consider two normalised
transition graphs
\[
G_i = ( N_i, \ii n_i, \Sigma, t_i : N_i\times\Sigma \pfun N_i, r_i : N_i \fun \mathbb{P}\mathbb{P}(\Sigma)),\qquad i = 1,2,
\]
over the same alphabet $\Sigma$. Their product is defined by
%
\begin{eqnarray}
G_1\times G_2 & = & (N_1\times N_2,(\ii n_1,\ii n_2), t:(N_1\times N_2)\times\Sigma\pfun (N_1\times N_2))
\\
\dom~t & = & \{ ((n_1,n_2),e)\in (N_1\times N_2)\times\Sigma~|   \nonumber
\\ & & \tabc
(n_1,e)\in\dom~t_1\wedge
(n_2,e) \in\dom~t_2    \}
\\
t((n_1,n_2),e) & = & (t_1(n_1,e),t_2(n_2,e))\ \text{for $((n_1,n_2),e)\in\dom~t$}
\end{eqnarray}
%
The following lemma is used in the proof of our main theorem.
%
\begin{lemma}\label{lemma:reachproduc}
If $G_1$ has $p$ states and $G_2$ has $q$ states, then every reachable state
$(n_1,n_2)$ of the product graph $G_1\times G_2$ can be reached by a trace
%%%$s\in\trc(G_1)\cap \trc(G_2)$  we don't need this
of maximal length $(pq-1)$.
\end{lemma}
\begin{proof}
The product graph $G_1\times G_2$ has at most $pq$ states. The empty trace $\varepsilon$
reaches its initial state $(\ii n_1,\ii n_2)$. Applying Lemma~\ref{lemma:extendV}
$(pq-1)$ times with $V=\{\varepsilon \}$ implies that $G_1\times G_2$ reaches
all of its reachable states (there are at most $pq$ of them) under
$V' = V \cup V.\Sigma\cup\dots \cup V.\Sigma^{(pq-1)}$. The maximal length of traces in
$V'$ is $(pq-1)$.
\xbox
\end{proof}

%\fixme{Abrupt end to the section. It has more than just preliminary material.
%It has some new results.}

>>>>

We present CSP~(Section~\ref{section:csp}) and the concept of minimal hitting
sets~(Section~\ref{sec:hit}), which is central to our novel notion of test
for failures refinement.

% =========================================================================
\subsection{CSP, Refinement, and Normalised Transition Graphs}
\label{section:csp}
% =========================================================================

@@

Throughout this paper, the alphabet of the processes, that is, the set of
events that are in scope, is denoted by $\Sigma$ and supposed to be finite.
The FDR tool~\cite{fdr} supports model checking and semantic analyses of
finite-state CSP processes.

@@

Two forms of choice support branching behaviour. An external choice $P
\extchoice Q$ between processes $P$ and $Q$ offers to the environment the
initial events of $P$ and $Q$. Once synchronisation takes place, the process
that has offered this event is chosen and the other is discarded. In an
internal choice $P \intchoice Q$, the environment does not have an
opportunity to interfere:~the choice is made by the process.
%
\begin{example}\label{example:CSP}
  We consider $P$, $Q$, and $R$ below. $P$ is initially
  ready to engage in the event $a$, and then makes an internal  choice to
  behave like either $Q$ or $R$.
  \begin{eqnarray*}
  P & = & a \then (Q\intchoice R)
  \\
  Q & = & a \then P \extchoice c \then P
  \\
  R & = & b \then P \extchoice c \then R
  \end{eqnarray*}
  $Q$, for instance, offers to the environment the choice to engage in $a$
  again or $c$. In both cases, afterwards, we have a recursion back to $P$.
  In $R$, if $b$ is chosen, we also have a recursion back to $P$. If $c$ is
  chosen, the recursion is to $R$.
  \xbox
\end{example}
%
Iterated forms $\Extchoice i: I @ P(i)$ and $\Intchoice i: I @ P(i)$ of the
external and internal choice operators define a choice over a collection of
processes $P(i)$. If the index set $I$ is empty, the external choice is the
process $\Stop$, which deadlocks:~does not engage into any event or
terminate. For an external choice $\Extchoice e: A @ e \then P(e), A\subseteq
\Sigma$ over a set $A$ of events, we use the abbreviation $e:A \then P(e)$.
An iterated internal choice is not defined for an empty index set.

@@

Interactions that are not supposed to be visible to the environment can be
hidden. The operator $P \hide H$ defines a process that behaves as $P$, with
the interactions modelled by events in the set $H$ hidden. Frequently, hiding
is used in conjunction with parallelism:~it is often desirable to make
actions of each process in a network of parallel processes, perhaps used for
coordination of the network, invisible, while events happening at their
interfaces remain observable.

@@

For divergence-free processes, failures refinement, $P \lessdet_F Q$, is
defined as expected by set containment $\fails(Q)\subseteq\fails(P)$. Since
refusals are subset-closed, $P \lessdet_F Q$ implies
$(s,\varnothing)\in\fails(P)$ for all traces $s\in\trc(Q)$. So, for
divergences-free processes, failures refinement implies trace refinement.
Therefore, using the conformance relation $conf$ below
%
\begin{equation}\label{eq:conf}
  Q\ conf\ P \defs \forall s\in \trc(P) \cap \trc(Q): \refs(Q/s)
  \subseteq \refs(P/s),
\end{equation}
%
failures refinement can be expressed by $\lessdet_T$ and $conf$.
%
\begin{equation}\label{thm:fref-tref-conf}
(P \lessdet_F Q) \iff (P \lessdet_T Q \land Q\ conf\ P)
\end{equation}
%
For finite CSP processes, since refusals are
subset-closed, $\refs(P/s)$ can be constructed from the set of \emph{maximal
refusals} defined by
%
\begin{equation}
\maxrefs(P/s) = \{ R \in\refs(P/s)~|~\forall R'\in \refs(P/s) - \{ R\}: R \not\subseteq R'\}
\end{equation}
%
Conversely, with the maximal refusals $\maxrefs(P/s)$ at hand, we can
reconstruct the refusals in the set $\refs(P/s)$ as described by
%
\begin{equation}\label{eq:refMaxRef}
\refs(P/s) = \{ R'\in\power(\Sigma)~|~\exists R\in \maxrefs(P/s): R'\subseteq R \}.
\end{equation}
%
Deterministic process states $P/s$ have exactly the one maximal refusal
defined by $\Sigma-[P/s]^0$, where $[P/s]^0$ denotes the \emph{initials} of
$P/s$, that is, the events that $P/s$ may engage into. The maximal refusals
in combination with the initials of a process express its degree of
nondeterminism.
%
\begin{example}
\label{ex:nondetdegree} $P = (\Stop \intchoice Q)$ has maximal refusals
$\maxrefs(P) = \{ \Sigma \}$, because $\Stop$ refuses to engage in any event,
and this is carried over to $P$ by the internal choice. However, $P$ is
distinguished from $\Stop$ by its initials, which are defined by $[P]^0 = [
\Stop \intchoice Q]^0 = [Q]^0$. So $P$ may engage nondeterministically in any
initial event of $Q$, but also refuse everything, due to internal selection
of $\Stop$.

@@

As shown in~\cite{Roscoe:1994:chapter}, the failures semantics of any
finite-state CSP process $P$ can be represented by a \emph{normalised
transition graph} $G(P)$ defined by a tuple
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$
with nodes $N$, initial node $\ii n\in N$, and process alphabet $\Sigma$. The
partial \emph{transition function} $t$ maps a node $n$ and an event
$e\in\Sigma$ to its successor node $t(n,e)$. If $(n,e)$ is in the domain of
$t$, then there is a transition, that is, an outgoing edge, from $n$ with
label $e$, leading to node $t(n,e)$. Normalisation of $G(P)$ is reflected by
the fact that $t$ is a function. The graph construction
in~\cite{Roscoe:1994:chapter} implies that all nodes $n$ in $N$ are
reachable by sequences of edges labelled by $e_1\dots e_k$ and connecting
states $\ii n,n_1,\dots,n_{k-1},n$, such that
\[
n_1 = t(\ii n,e_1), \quad n_i = t(n_{i-1},e_i),\ i = 2,\dots,k-1,\quad
n= t(n_{k-1},e_k).
\]
%
By construction, $s\in\Sigma^*$ is a trace of $P$, if, and only if, there is
a path through $G(P)$ starting  at $\ii n$ whose edge labels coincide with
$s$. In analogy to $\trc(P)$, we use the notation $\trc(G(P))$ for the set of
finite, initialised paths through $G(P)$, each path represented by its finite
sequence of edge labels. We note that $\trc(P) = \trc(G(P))$. Since $G(P)$ is
normalised, there is a unique node reached by following the events from $s$
one by one, starting in $\ii n$. Therefore, $G(P)/s$  is also well defined.
By $[n]^0$ we denote the \emph{initials} of $n$:~the set of events occurring
as labels in any outgoing transitions.
$$
[n]^0 = \{ e\in\Sigma~|~(n,e)\in\dom~t \}
$$
The graph construction from~\cite{Roscoe:1994:chapter} used to define
$G(P)$ for any process $P$ guarantees that $[G(P)/s]^0 = [P/s]^0$ for all
traces $s$ of $P$.

@@

When investigating  tests for failures refinement, the notion of
\emph{acceptances}~\cite{Hennessy:1988:ATP:50497}, which is dual to refusals,
is also useful. An acceptance set of $P/s$ is a subset of its initials
$[P/s]^0$; equivalently, it is a subset of events labelling  outgoing transitions of
$G(P)/s$. If the behaviour of  $P/s$ is deterministic, its only acceptance
equals $[P/s]^0$, because $P/s$ never refuses any of the events contained in
this set. If $P/s$ is nondeterministic, it internally chooses one of its
\emph{minimal acceptance sets} $A$ and never refuses any event in $A$, while
{\it possibly} refusing the events from $[P/s]^0 - A$ and
{\it always} refusing those from $\Sigma - [P/s]^0$.
The acceptances of $P/s$ are denoted by
$\accs(P/s)$, and the minimal acceptances by $\minaccs(P/s)$. They satisfy
the following properties.
%
\begin{eqnarray}
A\in \minaccs(P/s) & \Leftrightarrow & \exists R\in\maxrefs(P/s) \wedge A = \Sigma-R
\label{eq:accref}
\\
\bigcup \{ A~|~A\in \accs(P/s)\} & = & [P/s]^0
\label{eq:accinitials}
\\
 X\in\accs(P/s) &\Leftrightarrow & \exists A\in \minaccs(P/s): A\subseteq X \subseteq [P/s]^0
 \label{eq:accsubset}
\end{eqnarray}
%
%Note that (\ref{eq:accref}) can be regarded as a definition of minimal
%acceptances by means of maximal refusals. In this case,
%(\ref{eq:accinitials}) and (\ref{eq:accsubset}) are consequences of this
%definition and the fact that refusals are subset-closed. \fixme{alcc: are you
%sure? (4) does not talk about Acc. Is this important?}
%\fixme{jp: I'm sure and I can prove this. However, we don't need this fact, so I suggest to
%removed it.}
%
Exploiting (\ref{eq:accref}),
%(\ref{eq:accinitials}), and (\ref{eq:accsubset}),
every node of a normalised transition graph can
alternatively be labelled with their minimal acceptances, and this
information is equivalent to that contained in the maximal refusals. Since
process states $P/s$ are equivalently expressed by states $G(P)/s$ of $P$'s
normalised transition graph, we also write $\minaccs(G(P)/s)$ and
note that (\ref{eq:refsAndR}) and (\ref{eq:accref}) imply
\begin{equation}\label{eq:maxrefsminaccs}
\minaccs(G(P)/s) = \{ \Sigma - R | R\in r(G(P)/s) \}.
\end{equation}
\fixme{alcc: proof in the report perhaps? What do you need (5) for? }

% .....................................................................................
 \begin{figure}[t]
   %%\hspace*{-40mm}
   \begin{center}
     \includegraphics[width=\textwidth]{p.pdf}
   \end{center}
   %%\vspace*{-10mm}
   \caption{Normalised transition graph of CSP process $P$ from Example~\ref{ex:a}.}
   \label{fig:tga}
 \end{figure}

\fixme{alcc: the a from node 2 should go back to 0, I think.}

% .......................................................................................

@@

Having engaged in $c$, the next process state is represented by node 2. Due
to normalisation, there is only a single transition satisfying
$t(1,c) = 2$. This transition, however, can have
been caused by either $Q$ or $R$ engaging into $c$, so node 2 corresponds to
process state $Q/c \intchoice R/c = P \intchoice R$. This is reflected by the
two minimal acceptance sets labelling node 2. Similar considerations explain
the other nodes and transitions in Fig.~\ref{fig:tga}.
\xbox
\end{example}
%
Summarising, refinement relations between finite-state CSP processes $P, Q$ can be be
expressed by means of their normalised transition graphs
\begin{eqnarray*}
G(P) & = & ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))
\\
G(Q) & = & ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))
\end{eqnarray*}
%
as established by the results  in the following lemma.
There, result (\ref{eq:trcrefa}) reflects trace refinement in terms of graph
traces; (\ref{eq:failconf}) expresses failures refinement in terms of traces
refinement and $conf$; (\ref{eq:failrefa}) states how $conf$ can be expressed
by means of the maximal refusal functions of the graphs involved; and
(\ref{eq:failrefb}) states the same in terms of the minimal acceptances that
can be derived from the maximal refusal functions by means of
(\ref{eq:maxrefsminaccs}).
%%\fixme{alcc: I think at least the report should point to proofs.}
%
\begin{lemma}
  \label{lemma:tgtrcref}
  \begin{eqnarray}
  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
  \label{eq:trcrefa}
  \\
  \label{eq:failconf}
  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge Q\ conf\ P
  \\
  \label{eq:failrefa}
  Q\ conf\ P & \Leftrightarrow & \nonumber
  \forall s\in\trc(G(Q))\cap \trc(G(P)), R_Q\in r_Q(G(Q)/s):  \nonumber
  \\ & & \tabd
  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
  \\
  \label{eq:failrefb}
   & \Leftrightarrow &
    \forall s\in\trc(G(Q))\cap \trc(G(P)), A_Q\in\minaccs(G(Q)/s):  \nonumber
   \\ & & \tabd
  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
 \end{eqnarray}
  \xbox
\end{lemma}
%
\begin{proof}
Recall that $P\lessdet_T Q$ is defined as $\trc(Q)\subseteq \trc(P)$ and that
$\trc(P)=\trc(G(P))$ and $\trc(Q)=\trc(G(Q))$. This shows (\ref{eq:trcrefa}).
\fixme{alcc: (11) is already proved in my paper with mcg. No need to repeat
it here. I suggest just to add a reference.}

@@

For proving our main theorems, it is necessary to consider the \emph{product}
of normalised transition graphs. We need this only for the investigation of
corresponding traces in reference processes and processes for SUTs. So, the
labelling of nodes with maximal refusals or minimal acceptances are
disregarded in the product construction. We consider two normalised
transition graphs
\[
G_i = ( N_i, \ii n_i, \Sigma, t_i : N_i\times\Sigma \pfun N_i, r_i : N_i \fun \mathbb{P}\mathbb{P}(\Sigma)),\qquad i = 1,2,
\]
over the same alphabet $\Sigma$. Their product is defined by
%
\begin{eqnarray}
G_1\times G_2 & = & (N_1\times N_2,(\ii n_1,\ii n_2), t:(N_1\times N_2)\times\Sigma\pfun (N_1\times N_2))
\\
\dom~t & = & \{ ((n_1,n_2),e)\in (N_1\times N_2)\times\Sigma~|   \nonumber
\\ & & \tabc
(n_1,e)\in\dom~t_1\wedge
(n_2,e) \in\dom~t_2    \}
\\
t((n_1,n_2),e) & = & (t_1(n_1,e),t_2(n_2,e))\ \text{for $((n_1,n_2),e)\in\dom~t$}
\end{eqnarray}
%
The following lemma is used in the proof of our main theorem.
%
\begin{lemma}\label{lemma:reachproduc}
If $G_1$ has $p$ states and $G_2$ has $q$ states, then every reachable state
$(n_1,n_2)$ of the product graph $G_1\times G_2$ can be reached by a trace
%%%$s\in\trc(G_1)\cap \trc(G_2)$  we don't need this
of maximal length $(pq-1)$.
\end{lemma}
\begin{proof}
The product graph $G_1\times G_2$ has at most $pq$ states. The empty trace $\varepsilon$
reaches its initial state $(\ii n_1,\ii n_2)$. Applying Lemma~\ref{lemma:extendV}
$(pq-1)$ times with $V=\{\varepsilon \}$ implies that $G_1\times G_2$ reaches
all of its reachable states (there are at most $pq$ of them) under
$V' = V \cup V.\Sigma\cup\dots \cup V.\Sigma^{(pq-1)}$. The maximal length of traces in
$V'$ is $(pq-1)$.
\xbox
\end{proof}
%
This concludes our presentation of CSP and of some results regarding its
semantics and normalised transition graphs that are used in the next section.

\fixme{Abrupt end to the section. It has more than just preliminary material.
It has some new results.}
